---
title: "p8105_hw6_mh4588"
author: "Maggie Hsu"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse) #Load libraries
library(MASS) #load Mass library for variable selection
#libraries for cross validation
library(modelr)
library(mgcv)
library(SemiPar)

#for NOAA data
remotes::install_github("ropensci/rnoaa")

#Import data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  dplyr::select(name, id, everything())

homicide = read_csv("./homicide-data.csv") #Q2 homicide dataset
birthweight = read_csv("./birthweight.csv") #Q3 birthweight dataset

set.seed(8105) #set seed for randomization 
```
## Question 1
```{r weather}

```

## Question 2
```{r homicide}
homicide <- read_csv("./homicide-data.csv") #Import homicide dataset

#Clean data
homicide =  mutate(city_state = paste(city, state, sep=', '), homicide) |> #Create "city,state" variable
  filter(city_state != c("Dallas, TX","Phoenix, AZ","Kansas City, MO")) |> #Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO
   mutate(
     solved = case_when(
      disposition == c("Closed by arrest", "Closed without arrest") ~ 1, 
      disposition == "Open/No arrest" ~ 0)
     ) #Convert solved/unsolved to binary variable

homicide = filter(homicide, victim_race == "White"| victim_race == "Black") |> #Select victim_race to only have observations where victim_race is white or black
  mutate(victim_age = as.numeric(victim_age)) #"Unknown" age converted into NAs

#Omit data entry error "Tulsa-AL"
which(homicide == "Tul-000769", arr.ind=TRUE) #find row number
homicide <- homicide[-c(38414),]
```
```{r baltimore}
baltimore <- filter(homicide, city=="Baltimore") #Create a subset of the dataset homicide for just Baltimore observations.

baltimore_glm <- glm(solved~victim_age+victim_sex+victim_race, data=baltimore, family = binomial()) #Set up regression model

#Interpret results of regression model
baltimore_glm |>
  broom::tidy() |>
  mutate(OR = exp(estimate), 
         conf_lower = confint(baltimore_glm)[,1], #Confidence interval bounds
         conf_upper = confint(baltimore_glm)[,2]) |>
  filter(term == "victim_sexMale") |> #filter to adjusted odds ratio based on sex
  knitr::kable(digits = 3) 
```
```{r cities glm function}
#Define regression function 
city_glm = function(city) {
  city_glmdata =  filter(homicide, city_state == city)
  glm_city = glm(solved~solved~victim_age+victim_sex+victim_race, data=city_glmdata, family=binomial())
  
  glm_city |>
  broom::tidy() |>
  mutate(OR = exp(estimate), 
         conf_lower = confint(glm_city)[,1],
         conf_upper = confint(glm_city)[,2]) |>
  filter(term == "victim_sexMale") |>
  knitr::kable(digits = 3) 
}

```

```{r cities glm iteration}
homicide_cities = distinct(homicide["city_state"]) #list of cities to iterate over
output = vector("list", length = nrow(homicide_cities)) #output list

purrr::map(homicide_cities, city_glm)
```
## Question 3
```{r birthweight}
#Model with birth length and gestational age
birthw_lengthage <- lm(bwt~blength+gaweeks, data=birthweight)
summary(birthw_lengthage)
#Model with head circumference, length, sex, and interactions
birthw_interactions <- lm(bwt~bhead+blength+babysex+bhead*blength+blength*babysex+bhead*babysex+bhead*babysex*blength, data=birthweight)
summary(birthw_interactions)
```
```{r bwt design}
#Proposed design model
#Variable selection using stepwise criteria and AIC/BIC criterion
birthw_full <- lm(bwt~.,data=birthweight)
birthw_stepwise <- stepAIC(birthw_full, direction = "both", trace = FALSE, k=2)
birthw_stepwise_bic <- stepAIC(birthw_full, direction = "both", trace = FALSE, k=log(4342))

#model selection
birthw_aic <- lm(bwt ~ babysex + bhead + blength + delwt + fincome + 
    gaweeks + menarche + mheight + momage + mrace + parity + 
    ppwt + smoken, data = birthweight)
summary(birthw_aic)

birthw_bic <- lm(bwt ~ babysex + bhead + blength + delwt + fincome + 
    gaweeks + mrace + ppwt + smoken, data = birthweight)
summary(birthw_bic)

```
I used the AIC and BIC criteria to select variables from the full model based on the lowest AIC or BIC value. Since the BIC-minimized model has a higher R-squared but less variables than the AIC-minimized model (less overfitting), I would go with the BIC model for this problem.
```{r fitted residual}
birthweight = add_predictions(birthweight, birthw_bic) #add fitted values
birthweight = add_residuals(birthweight, birthw_bic) #add residual values

#fitted vs. residual plot
ggplot(data=birthweight)+aes(x=pred, y=resid)+geom_point() + labs(title="Fitted vs. Residual Values")+xlab("Fitted Values")+ylab("Residual Values")
```
My proposed model has a higher R-squared and adjusted R-squared than the two previous models birthw_lengthage and birthw_interactions